---
title: "Global Greenhouse Gas Emissions"
subtitle: "Data Visualisation Exercise"
author: "Valentin Klotzb&uuml;cher"
authors: 
  - name: Valentin Klotzbücher
    orcid: 0000-0001-9382-6757
    email: valentin.klotzbuecher@econ.uni-freiburg.de
#     affiliations:
#       - ref: uf
# affiliations:
#    - id: uf 
#      name: University of Freiburg 
#      address: Freiburg im Breisgau, DE
date: last-modified
date-format: long
bibliography: 
  - "references.bib"
  - "grateful-refs.bib"
csl: "nature.csl"
link-citations: true
crossref:
  fig-title: "Chart"
format:
  pdf: 
    papersize: a4
    toc: false
    lof: false
    lot: false
    colorlinks: true
    geometry:
      - top = 2cm
      - right = 2cm
      - bottom = 2cm
      - left = 2cm
    mainfont: "Arial"
    sansfont: "Arial"
    monofont: "Arial"
    execute: 
      echo: false
      warning: false
      message: false
      cache: false
    fig-cap-location: top
    fig-align: left
    fig-width: 10
    fig-height: 6
    include-in-header: include-in-header.tex
    template-partials:
        - before-body.tex
  html:
    theme: [journal, theme-light.scss]
    highlight-style: atom-one
    toc: false
    citations-hover: true
    footnotes-hover: true
    callout-appearance: minimal
    embed-resources: true
    smooth-scroll: true
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    code-overflow: wrap
    fig-cap-location: top
    fig-align: left
    lightbox: true
    execute: 
      warning: false
      message: false
    code-links:
      - text: Replication files
        icon: github
        href: https://github.com/valentinklotzbuecher/ECB_2025
    other-links: 
      - text: My website
        icon: bullseye
        href: https://valentink.quarto.pub/
---

This report analyzes global greenhouse gas (GHG) emissions using data from the Emissions Database for Global Atmospheric Research (EDGAR) GHG emissions files [@edgar2024; @edgar2024data]. Information on countries' income-groups was obtained from the World Bank API [@wbstats].

::: {.content-hidden unless-format="html"}
The analysis was conducted in R version 4.4.2 [@base], using a number of additional packages for data preparation and visualisation [@countrycode; @here; @janitor; @renv; @rmarkdown2024; @scales; @showtext; @tidyverse; @treemapify; @zoo], and compiled into a reproducible PDF report with Quarto.
:::

::: {.content-hidden unless-format="pdf"}
The analysis was conducted in R version 4.4.2 [@base], using a number of additional packages for data preparation and visualisation [@countrycode; @grateful; @here; @janitor; @rmarkdown2024; @scales; @showtext; @tidyverse; @treemapify; @zoo], and compiled into a reproducible PDF report with Quarto. The code to reproduce this PDF document, together with an HTML version and all required files, is available at [valentink.quarto.pub/ghg](https://valentink.quarto.pub/ghg/).
:::

::: {#fig-1}

```{r}
#| label: chart1

library("tidyverse")
library("janitor")
library("here")
library("readxl")
library("scales")
library("zoo") 
library("treemapify")

library("showtext")
font_add("Arial", regular = "arial.ttf")
showtext_auto()

library("wbstats")
country_meta <- wb_countries() |>
  mutate(country_wb = country) |>
  select(iso3c, country_wb, region, income_level)

library("countrycode")

library("grateful")
# citations <- cite_packages(output = "paragraph", out.dir = ".", cite.tidyverse = TRUE)

euro_area_countries <- c(
  "Austria", "Belgium", "Croatia", "Cyprus", "Estonia", "Finland", "France and Monaco",
  "Germany", "Greece", "Ireland", "Italy, San Marino and the Holy See", "Latvia", "Lithuania",
  "Luxembourg", "Malta", "Netherlands", "Portugal", "Slovakia", "Slovenia", "Spain and Andorra"
)
 
edgar_growth <- read_excel(here("data", "EDGAR_2024_GHG_booklet_2024.xlsx"), sheet = "GHG_totals_by_country") |>
  clean_names()  |>
  filter(!is.na(country)) |>
  pivot_longer(cols = starts_with("x"), names_to = "year", values_to = "emissions") |>
  mutate(
    year = as.numeric(str_remove(year, "x")),
    emissions = ifelse(emissions == 0, NA, emissions),
    euro_area = if_else(country %in% euro_area_countries, 1, 0)
  )  |>
  filter(!country %in% c("International Aviation", "International Shipping"))


edata_euroarea <- edgar_growth |>
  filter(country %in% euro_area_countries) |>
  group_by(year) |>
  summarise(emissions = sum(emissions, na.rm = TRUE)) |>
  arrange(year) |>
  mutate(echange_euroarea = emissions - lag(emissions)) |>
  mutate(egrowth_euroarea = (emissions - lag(emissions))/lag(emissions)) |>
  mutate(egrowth_euroarea_MA = zoo::rollmean(egrowth_euroarea, k = 5, fill = NA, align = "right")) |> 
  select(year, starts_with("egrowth")) |>
  mutate(region = "EA") 

edata_eu27 <- edgar_growth |>
  filter(country == "EU27") |>
  arrange(year) |>
  mutate(echange_eu27 = emissions - lag(emissions)) |>
  mutate(egrowth_eu27 = (emissions - lag(emissions))/lag(emissions)) |> 
  mutate(egrowth_eu27_MA = zoo::rollmean(egrowth_eu27, k = 5, fill = NA, align = "right")) |> 
  select(year, starts_with("egrowth")) |>
  mutate(region = "EU27") 
  
edata_world <- edgar_growth |>
  filter(country == "GLOBAL TOTAL") |>
  arrange(year) |>
  mutate(echange_world = emissions - lag(emissions)) |>
  mutate(egrowth_world = (emissions - lag(emissions))/lag(emissions)) |>
  mutate(egrowth_world_MA = zoo::rollmean(egrowth_world, k = 5, fill = NA, align = "right")) |> 
  select(year, starts_with("egrowth")) |>
  mutate(region = "World") 

edata_regions <- edata_world |>
  left_join(edata_eu27, by = "year") |>
  left_join(edata_euroarea, by = "year")  |>
  filter(year != 1970 ) |>
  select(year, starts_with("egrowth")) |>
  pivot_longer(cols = -year, names_to = "series", values_to = "egrowth") |>
  mutate(
    region = case_when(
      str_detect(series, "world") ~ "Worldwide",
      str_detect(series, "eu27") ~ "EU 27",
      str_detect(series, "euroarea") ~ "Euro Area"
    ),
    type = if_else(str_detect(series, "_MA$"), "MA", "value")
    ) |>
  mutate(region = factor(region, levels = c("Euro Area", "EU 27", "Worldwide"))) |>
  pivot_wider(names_from = type, values_from = egrowth)  |>
  group_by(year, region) |>
  summarise(
    value = first(value[!is.na(value)]),
    MA = first(MA[!is.na(MA)]),
    .groups = "drop"
  ) |>
  mutate(tval = round(value*100, 1),
         tMA = round(value*100, 1)) |>
  filter(year>1974)

chart1 <- ggplot(edata_regions, aes(x = year, color = region)) +
  geom_segment(aes(x = year, xend = year, y = 0, yend = value), 
               color = "grey90", size = 0.5) +
  geom_point(aes(y = value), size = 2, alpha = 0.7) +
  geom_line(aes(y = MA), size = 1, alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.5) +
   scale_color_manual(values = c(
    "Euro Area" = "#003299",
    "EU 27" = "#ffb400",
    "Worldwide" = "#ff4b00"
  )) +
  scale_y_continuous(labels = scales::percent_format(), position = "left") +
  labs(
    x = element_blank(),
    y = "GHG Emissions, Growth Rate",
    color = element_blank()
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", color = "black", size = 17),
    axis.text = element_text(color = "black", size = 15),
    panel.grid.major = element_line(size = 0.4, linetype = 2),
    panel.grid.minor = element_blank(),
    legend.position = "top",
    legend.justification = c(0, 1),
    legend.key.size = unit(1, "lines")
  )

chart1



```

*Notes:* Growth rate of GHG emissions over time, in percent. Each lines shows the 5-year leading moving average, colored by region, with points indicating the underlying year-on-year growth rates. GHG emissions include CO2 (fossil only), CH4, N2O and F-gases. \newline 
*Sources:* EDGAR @edgar2024data and own calculations.

**Evolution of GHG growth in the euro area, European Union (EU27) and worldwide.**

:::

*Main insights:*

-   Global deceleration, but still growing: Over the last 15 years, GHG emissions growth rates have decelerated across all regions. However, global emissions grew by `r edata_regions |> filter(region == "Worldwide" & year == "2023") |> pull(tval)`% in 2023 and the global 5-year moving average (MA) remains positive throughout the period, reflecting continued emissions growth worldwide.

-   Negative growth in EU and EA: The 5-year moving average for emissions growth has been negative in the EU and EA since `r edata_regions |> filter(region %in% c("EU 27", "Euro Area") & MA < 0) |> summarise(min(year)) |> pull()`, for 2023 we observe a change of `r edata_regions |> filter(region == "Worldwide" & year == "2023") |> pull(tval)`%.

-   Impact of global crises: Significant global reductions were observed only during major crises, such as the global financial crisis and COVID-19 pandemic. In 2020, GHG emissions decreased by `r edata_regions |> filter(region == "Worldwide" & year == "2020") |> pull(tval)`% worldwide and even stronger in the European Union (`r edata_regions |> filter(region == "EU 27" & year == "2020") |> pull(tval)`% and the Euro Area (`r edata_regions |> filter(region == "Euro Area" & year == "2020") |> pull(tval)`% ).

\newpage

::: {#fig-2}

```{r}
#| label: chart2

edgar_pc <- read_excel(here("data", "EDGAR_2024_GHG_booklet_2024.xlsx"), sheet = "GHG_per_capita_by_country") |>
  clean_names() |>
  filter(country != "GLOBAL TOTAL" 
         & country != "EU27" 
         & country != "") |>
  pivot_longer(cols = starts_with("x"), names_to = "year", values_to = "emissions") |>
  mutate(year = as.numeric(str_remove(year, "x"))) |>
  filter(year == max(year)) |>
  left_join(country_meta, by = c("edgar_country_code" = "iso3c")) |>
  mutate(
    income_level = if_else(
      is.na(income_level), "Not classified",income_level)
  )

edgar_pc <- edgar_pc |>
  mutate(lemissions = log(emissions)) |> 
  mutate(income_level = factor(income_level, 
                               levels = c("Not classified",
                                          "Low income", 
                                          "Lower middle income",
                                          "Upper middle income",
                                          "High income")))

 group_means <- edgar_pc |>
  group_by(income_level) |>
  summarise(med_lemissions = median(lemissions, na.rm = TRUE),
            med_emissions = median(emissions, na.rm = TRUE))

edgar_pc <- edgar_pc |>
  mutate(
    label = case_when(
      country %in% c("Venezuela", "France", "China", "Nigeria") ~ country,  # Label selected countries
      # emissions < quantile(emissions, 0.05, na.rm = TRUE) ~ country,  # Label outliers
      TRUE ~ NA_character_  # No label for other countries
    )
  )

# Plot
chart2 <- ggplot(edgar_pc, aes(x = income_level, y = emissions)) +
  geom_jitter(aes(color = income_level), width = 0.35, alpha = 0.6, size = 5) +
  geom_point(data = group_means, aes(x = income_level, y = med_emissions),
             shape = 21, size = 4, fill = "white", color = "black", stroke = 1, alpha = 1) +
  # geom_text_repel(aes(label = label, color = income_level), size = 3, show.legend = FALSE) +  
  # geom_text(data = group_means, aes(x = income_level, y = med_emissions, label = scales::comma(med_emissions)),
            # vjust = -0.5, size = 3, color = "black") +
  scale_color_manual(values = c(
    "Low income" ="#65b800",
    "Lower middle income" ="#ff4b00",
    "High income" =  "#003299",
    "Upper middle income" =  "#ffb400",
    "Not classified" = "gray60"
  )) +
  scale_y_continuous(
    trans = "log",                             
    breaks = c(1, 2, 4, 8, 16, 32, 64),  
    position = "left"                          
  ) +
  labs(
    x = element_blank(),
    y = "Emissions per capita in t CO2eq/person (log scale)"
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", color = "black", size = 17),
    axis.text = element_text(color = "black", size = 15),
    legend.position = "none", 
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_line(linewidth = 0.4, linetype = 2)
  ) +
  coord_flip()

chart2

```

*Notes:* Emissions per capita in 2023, in tons of CO2eq/person (log scale). Individual countries are sorted and colored by income group, with black circles denoting the median value for each group. GHG emissions include CO2 (fossil only), CH4, N2O and F-gases. 
\newline
*Sources:* EDGAR @edgar2024data, World Bank @wbstats, and own calculations.

**Comparison of countries’ GHG emissions per capita aggregated according to the World Bank income groups.**

:::

*Main insights:*

-   Per capita emissions across income groups: Median per capita emissions increase steeply with income, from `r group_means |> filter(income_level == "Low income") |> pull(med_emissions) |> round(2)` t CO2-eq per person in low-income countries, to `r group_means |> filter(income_level == "Lower middle income") |> pull(med_emissions) |> round(2)` in lower-middle income, `r group_means |> filter(income_level == "Upper middle income") |> pull(med_emissions) |> round(2)` in upper-middle income, and `r group_means |> filter(income_level == "High income") |> pull(med_emissions) |> round(2)` in high-income countries.

-   Stark differences: All low and lower-middle-income countries fall below the high-income median of `r group_means |> filter(income_level == "High income") |> pull(med_emissions) |> round(2)` t CO2-eq per person and -- except for one outlier -- per capita emissions across high-income countries are above the median of the lower-middle income group (`r group_means |> filter(income_level == "High income") |> pull(med_emissions) |> round(2)` t CO2-eq per person).

-   Variation increasing with income: Upper-middle and high-income groups exhibit broader within-variation in emissions.

\newpage

::: {#fig-3}

```{r}
#| label: chart3

edgar_cont <- read_excel(here("data", "EDGAR_2024_GHG_booklet_2024.xlsx"), sheet = "GHG_totals_by_country") |>
  clean_names() |>
  filter(!country %in% c("GLOBAL TOTAL", "EU27", "International Aviation", "International Shipping")) |>  
  pivot_longer(cols = starts_with("x"), names_to = "year", values_to = "emissions") |>
  mutate(year = as.numeric(str_remove(year, "x"))) |>  
  filter(!is.na(country)) |>
  filter(year == max(year)) 

edgar_cont <- edgar_cont |>
  mutate(
    continent = countrycode(sourcevar = edgar_country_code, 
                            origin = "iso3c",               
                            destination = "continent")
    ) |>   
  mutate(continent = case_when(country %in% c("Curaçao") ~ "Americas",
                               country %in% c("Serbia and Montenegro") ~ "Europe",
                               is.na(country) ~ "NA",
                               .default = continent))

edgar_cont <- edgar_cont |>
  arrange(desc(emissions)) |>
  mutate(label = if_else(row_number() <= 15, country, edgar_country_code))   |>   
  mutate(label = if_else(country %in% c("Türkiye","Saudi Arabia", "South Korea", "Australia"), edgar_country_code, label))

continent_colors <- c(
  "Africa" = "#00b1ea",
  "Asia" = "#003299",
  "Europe" = "#ffb400",
  "Americas" = "#ff4b00",
  "Oceania" = "#65b800"
)
 
chart3 <- ggplot(edgar_cont, aes(area = emissions, fill = continent, subgroup = continent, label = label)) +
  geom_treemap(color = "white") +
  geom_treemap_subgroup_border(color = "white", size = 1.5) + 
  geom_treemap_text(fontface = "italic", colour = "white", place = "centre", grow = TRUE) +  
  scale_fill_manual(values = continent_colors) +
  labs(fill = element_blank()) +
  theme_minimal() +
  theme(
    text = element_text(family = "Arial", color = "black", size = 17),
    axis.text = element_text(color = "black", size = 15),
    legend.position = "top",
    legend.justification = c(0, 1),
    legend.key.size = unit(0.7, "lines"),
    )

chart3

```

*Notes:* Contributions of countries to total global GHG emissions in 2023. The size of each rectangle, relative to the full area, reflects each country's contribution, sorted and colored by continent. 
\newline 
*Sources:*  EDGAR @edgar2024data and own calculations.

**Contribution of individual countries and continents to total world GHG emissions.**

:::

*Main insights:*

-   Continents: Asia is the largest contributor to global GHG emissions in 2023, accounting for `r edgar_cont |> filter(continent == "Asia") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%, followed by the Americas (`r edgar_cont |> filter(continent == "Americas") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%), Europe (`r edgar_cont |> filter(continent == "Europe") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%), and Africa (`r edgar_cont |> filter(continent == "Africa") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%). Oceania contributes the least at `r edgar_cont |> filter(continent == "Oceania") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%.

-   Country-level contributions: Individual countries play a disproportionate role, with China contributing `r edgar_cont |> filter(country == "China") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%, followed by the USA (`r edgar_cont |> filter(country == "United States") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%), India (`r edgar_cont |> filter(country == "India") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%), Russia (`r edgar_cont |> filter(country == "Russia") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%), and Brazil (`r edgar_cont |> filter(country == "Brazil") |> summarise(percent = sum(emissions, na.rm = TRUE) / sum(edgar_cont$emissions, na.rm = TRUE) * 100) |> pull(percent) |> round(1)`%).

-   Political implications: Decisions by the top-emitting countries, especially China, the USA, and India, are crucial for any effective global mitigation strategy. Coordinated international policies targeting high-emission regions could lead to significant reductions, given their outsized impact on total emissions.

\vfill

------------------------------------------------------------------------

\footnotesize

::: {#refs}
:::
